Discovering Focused Microgenre Communities

### 2023-08-03 main.tex
Updates from Overleaf

\subsection{Heavy Metal}
Heavy Metal is one of the most storied macrogenre labels in the sociology of taste. It is featured in Bethany Bryson's \citeyearpar{bryson96} classic paper. Since then, it has been the subject of scholarly debate regarding its possibly changing status in the cultural stratification order. Heavy Metal was distinct in Bryson's study for several reasons. First, it was the most disliked genre in the 1993 General Social Survey culture module data---one of the most heavily analyzed data sources in the sociology of taste, hence its being featured in the title of Bryson's eponymous piece. Second, it was the prototypical genre that was relatively more favored by working-class audiences but absolutely hated by individuals with high education. Thus, it served as the linchpin of Bryson's symbolic exclusion argument; omnivores were omnivorous up to a point, except when it came to those genres most favored by those with low education.

Regardless, the picture we got of Metal concerning audience segmentation was clear: It was a genre that primarily appealed to working-class, low-education white men and was rejected by pretty much everyone else. Lizardo and Skiles \citeyearpar[][6, table 2]{lizardo_skiles16} used the same data source as in this paper (data from 2012) and found that, nineteen years later, Metal was still the most disliked of the macrogenre labels; not only that, when asked who the typical fan of metal looked like, Americans agreed with the stereotype: Low education, ``lower'' and working-class white men \citep[][7, table 3]{lizardo_skiles16}.

Yet, cracks began to appear in this picture as soon as it was hung. First, Tampubolon \citeyearpar{tampubolon2008revisiting} re-analyzed the Bryson data using a different methodological approach (latent-class analysis), making different inclusion decisions regarding the ``don't know'' responses. Tampubolon shows that including those ``don't know'' responses significantly changes things. Concerning Metal, it indicates that it is strongly disliked among low-education more than would have been surmised in Bryson's original study. For Tampubolon, this makes Metal ``quite exceptional'' since it is disliked across the board and even more by low-education people.

So just because Metal's audience is composed of low-education people does not mean that the rest of the less-educated population loves it; in fact, they hate it passionately. Tampubolon's analysis of ``Like'' responses also shows that Metal, rather than being exclusively preferred by a univorous group of ``Metalheads'' (who only prefer metal and dislike everything else), also makes it into the taste profile of one of two omnivorous groups uncovered in his analysis. Thus, Metal ends up straddling omnivores and univores. Because of this, even in the 1993 GSS data, its association with markers of status was ambiguous, leading Tampubolon to conclude that ``[p]reference or dislike for heavy metal is, therefore, orthogonal to status as measured by education'' \citeyearpar[][257]{tampubolon2008revisiting}, and that ``heavy metal is definitely not a low culture in the sense of culture `strongly associated' or very much liked by those with low education.''

Tampubolon's argument regarding Metal dovetail with those preferred by Lizardo and Skiles \citeyearpar{lizardo_skiles15}, using the same data source as in this paper. Yes, while Metal was still the most disliked genre in 2012, \textit{compared} to 1993, it was one of the macrogenre labels (Rap and Hip Hop being the other) that experienced the steepest \textit{declines} in being disliked in the intervening two decades, especially among younger audiences. This strongly suggests that enjoying some types of Metal music may be an ``emerging form'' of cultural capital \cite{prieur2013emerging} among certain segments of the young elite.

\subsubsection{Microgenre Analysis}
So, will the real heavy metal please stand up? Is Metal a form of emerging cultural capital with an appeal to elites or a genre heavily favored by young white people with low education? Or is that an outdated picture of the American musical field? Is Metal even correlated with status markers like education at all, as Tampubolon suggested? Is speaking of Metal as a ``youth'' music accurate, or has its audience been aging along with the genre (going on more than four decades and counting)? Given prior work, a strong expectation is that, at the very least, we should observe a microgenre partition separating the ``prototypical'' Metal (appealing to low-education white men) and perhaps a less-prototypical one, appealing to younger audiences with high-level cultural capital with other micro-variations, perhaps track generational distinction \citep{lizardo_skiles15}.

To examine these questions, I partitioned the Heavy-Metal branch of the dendrogram, at a height that resulted in four microgenre variations. To explore the issue of audience segmentation, I specify a logistic regression model with five binary outcomes: First, a binary variable indicating that a person chose (liked and listened to) the standard macrogenre label, and then four models with the microgenre variations as the outcome. The results are shown in Table~\ref{tab:metal}.

>>>>

\subsection{Heavy Metal}
Heavy Metal is one of the most storied macrogenre labels in the sociology of taste, as featured in the title of Bethany Bryson's \citeyearpar{bryson96} now classic paper. Since then, it has been the subject of scholarly debate regarding its possibly changing status in the cultural stratification order \citep{tampubolon2008revisiting, goldberg2011mapping, lizardo_skiles15}. Heavy Metal stood out in Bryson's study for several reasons. First, it was the most disliked genre in the 1993 General Social Survey culture module data---one of the most heavily analyzed data sources in the sociology of taste---hence the ``Anything But'' in the title of Bryson's eponymous piece. Second, it was the prototypical genre that was relatively more favored by working-class audiences but absolutely hated by individuals with high education. Thus, it served as the linchpin of Bryson's symbolic exclusion argument; omnivores were omnivorous up to a point, except when it came to those genres most favored by those with low education.

Regardless, the picture we got of Metal concerning audience segmentation was clear: It was a genre that primarily appealed to working-class, low-education white men and was rejected by pretty much everyone else. Lizardo and Skiles \citeyearpar[][6, table 2]{lizardo_skiles16} used the same data source as in this paper (data from 2012) and found that, nineteen years later, Metal was still the most disliked of the macrogenre labels in a representative sample of Americans. Not only that, but when asked who the typical fan of metal looked like, Americans agreed with the stereotype: Low education, ``lower'' and working-class white men \citep[][7, table 3]{lizardo_skiles16}.

Yet, cracks began to appear in this picture as soon as it was hung. First, Tampubolon \citeyearpar{tampubolon2008revisiting} re-analyzed the Bryson data using a different methodological approach (latent-class analysis), making different inclusion decisions regarding the ``don't know'' responses. Tampubolon shows that including those ``don't know'' responses significantly changes things. Concerning Metal, it indicates that it is strongly disliked among low-education people more than would have been surmised in Bryson's original study, so it is not just a top-down symbolic exclusion story. For Tampubolon, this makes Metal ``quite exceptional'' since it is disliked across the board and even more by low-education people.

Tampubolon's work indicates that just because Metal's audience is composed of low-education people does not mean that the rest of the less-educated population loves it; in fact, they hate it passionately. Tampubolon's analysis of ``Like'' responses also shows that Metal, rather than being exclusively preferred by a univorous group of ``Metalheads'' (who only prefer metal and dislike everything else), also makes it into the taste profile of one of two omnivorous groups uncovered in his analysis. Thus, Metal ends up straddling omnivores and univores. Because of this, even in the 1993 GSS data, its association with markers of status was ambiguous, leading Tampubolon to conclude that ``[p]reference or dislike for heavy metal is, therefore, orthogonal to status as measured by education'' \citeyearpar[][257]{tampubolon2008revisiting}, and that ``heavy metal is definitely not a low culture in the sense of culture `strongly associated' or very much liked by those with low education'' (\textit{ibid}).

Tampubolon's argument regarding Metal dovetail with those preferred by Lizardo and Skiles \citeyearpar{lizardo_skiles15}, using the same data source as in this paper. Yes, while Metal was still the most disliked genre in 2012, \textit{compared} to 1993, it was one of the macrogenre labels (Rap and Hip Hop being the other) that experienced the steepest \textit{declines} in being disliked in the intervening two decades, especially among younger audiences. This led Lizardo and Skiles \citeyearpar[][18]{lizardo_skiles15} to conclude that ``The fact that young, highly educated Americans are now about as equally unlikely to dislike\ldots Heavy Metal as their same-age non-college educated peers constitutes a dramatic reversal of the pattern noted by Bryson\ldots [Metal] has experienced an improvement in standing among the college educated across almost all levels of age.'' This strongly suggests that enjoying some types of Metal music may be an ``emerging form'' of cultural capital \cite{prieur2013emerging} among certain segments of the young elite.

\subsubsection{Microgenre Analysis}
So, will the real heavy metal please stand up? Is Metal a form of emerging cultural capital with potential appeal to young middle-class people or a genre heavily favored by white men with low education? Or is that an outdated picture of the American musical field? Is Metal even correlated with status markers like education at all, as Tampubolon suggested? Is speaking of Metal as a ``youth'' music accurate, or has its audience been aging along with the genre (going on more than four decades and counting)? Given prior work, a strong expectation is that, at the very least, we should observe a microgenre partition separating the ``prototypical'' Metal (appealing to low-education white men) and perhaps a less-prototypical one, appealing to younger audiences with high-level cultural capital with other micro-variations, perhaps track generational distinction \citep{lizardo_skiles15}.

\begin{figure}[ht!]
    \centering    
    \includegraphics[width=1.3\textwidth]{Tabs/reg-tab-metal.png}
    \caption{}
    \label{tab:reg-metal}
\end{figure}

To examine these questions, I partitioned the Heavy-Metal branch of the dendrogram, at a height that resulted in four microgenre variations. To explore the issue of audience segmentation, I specify a linear probability regression model with five outcomes: First, a binary variable indicating that a person chose (liked and listened to) the standard macrogenre label (Heavy Metal), and then four models with the microgenre variations as the outcome (Heavy Metal one through four). The results are shown in Table~\ref{tab:reg-metal}. The main predictors are the respondent's education, an indicator variable for whether at least one of the respondent's parents has a college degree, respondent's age, racial identification (in five categories), and gender identification (in two categories). Because the age effect is specified using a squared polynomial, it can be hard to interpret its substantive significance. To help with this, I show the predicted probabilities obtained for the model at different values of age for the four different microgenres in Figure~\ref{fig:age-metal}.

\begin{figure}[ht!]
    \centering    \includegraphics[width=1.0\textwidth]{Plots/Demog/micro-by-age-metal.png}
    \caption{}
    \label{fig:age-metal}
\end{figure}

As shown in the first model, the socio-demographic predictors of choosing the ``Heavy Metal'' macrogenre label are the ones we would expect from Bryson's original work and the stereotypical perception of the genre's audience held by Americans. Metal fans are less likely to have high levels of education, are less likely to be older, are more likely to be white (negative effects for all other racial identifications compared to the base category), and are more likely to be men ($p<0.01$). So much for the emerging cultural capital story?

Not so fast. The results are different when looking at the predictors of the microgenre variation choices. Indeed, the predictors of choosing the first microgenre variation, shown in the second column of the table, are consistent with Tampubolon's story. For this microgenre, there is a flat education gradient ($p = 0.54$), and importantly, it is more likely to be consumed by people whose parents have a college degree ($p < 0.05$), indicating a link to cultural capital in the home environment. This microgenre is tilted towards younger audiences, as only the linear negative age term is statistically significant (see Figure~\ref{fig:age-metal}) and is also neutral with respect to race and gender. This version of Metal appeals mainly to young middle-class respondents and does not repel women or non-white audiences, a stark difference from what we would have concluded from observing the macrogenre label effects.

Note that the other microgenre variations conform most closely to the audience segmentation we would expect from the version of Metal that would appeal to ``univore'' consumers, indicating that the link clustering can be discerned between these variants. Particularly, the second microgenre, whose predictors are shown in the third column, seems to be the ``uber-Metal,'' at least regarding the stereotype. It has a negative education gradient ($p = 0.06$), appeals to young-adult and middle-aged audiences (see Figure~\ref{fig:age-metal}), and repels Black and Hispanic people as well as women. The other Metal microgenres are variations on the same theme: The third microgenre, as shown in the fourth column in the table, is like the second but appeals mainly to young audiences, and the fourth Metal microgenre is also similar to the second but appeals to a comparatively older crowd. So the remaining microgenre partitions seem to indicate variations within the ``univore'' consumption profile, perhaps indicative of distinctive versions of Metal with an appeal to different generational niches \citep{koch2020evolutionary}.

\subsection{Salsa}
``it is plausible that increasing appreciation among high-status groups for a genre conventionally defined as ‘popular’ or ‘lowbrow’, will go hand in hand with its symbolic upgrading. And when a genre expands and spreads in terms of production, distribution and reception, it is likely to differentiate into distinct styles or subgenres. Under these conditions, it is to be expected that differential appreciations for different styles and forms within a given genre will develop, which may end up in a more or less clear hierarchy within this genre'' \citeyearpar[][62]{Bachmayer2014-pk}

\begin{figure}[ht!]
    \centering    
    \includegraphics[width=1.3\textwidth]{Tabs/reg-tab-latin.png}
    \caption{}
    \label{tab:reg-metal}
\end{figure}

\begin{figure}[ht!]
    \centering    \includegraphics[width=1.0\textwidth]{Plots/Demog/micro-by-age-latin.png}
    \caption{}
    \label{fig:age-salsa}
\end{figure}
